% ================================== HEADER ====================================
\documentclass{article}           % Sets style/look of many things.
% \documentclass{report}          % part, chapters, front page etc.
\usepackage{exsheets}
\usepackage[utf8]{inputenc}       % Encoding of input files UTF-8
\usepackage[T1]{fontenc}
\usepackage[scaled]{beramono}     % Font
\usepackage{color}                % Color text
\usepackage{titlesec}             % Select alternative section titles
\usepackage{fancyvrb}
\usepackage{verbatim}             % Comment environment
\usepackage{listings}             % Format and render text/code etc.
\usepackage{minted}               % Much better syntax highlighting
\usepackage{float}                % Control of floating environment/figure
\usepackage{graphicx,  subfigure} % Better figures, graphics, units etc.
\usepackage{multicol}             % Multiple columns
\usepackage{amsmath}              % Math: Equation, split, align etc.
\usepackage{siunitx}              % SI units
\usepackage{mathtools}            % Different math tools to use with amsmath
\usepackage{amssymb}              % Math symbols
\usepackage[
    colorlinks,
    citecolor=black,              % I like links with standard black color
    filecolor=black,
    linkcolor=black,
    urlcolor=black
]{hyperref}                       % Links in TOC etc.
\usepackage[all]{hypcap}          % Better links to floating environment

\usepackage{tabto}
\newcommand\marginsymbol[1][0pt]{%
    \tabto*{0cm}\makebox[\dimexpr-1cm-#1\relax][r]{$\mathbb{P}$}\tabto*{\TabPrevPos}}

\renewcommand{\thesubsection}{\thesection.\alph{subsection}}
\title{\vspace{-2cm}INF3490/INF4490 Exercises - Week 1}
\author{Eivind Samuelsen, Ole Herman S. Elgesem}
\date{\today}

% Removing paragraph indents is sometimes useful:
\setlength\parindent{0pt}

% Make margins smaller to fit more figures, tables etc on page: (optional)
\addtolength{\oddsidemargin}{-1.0in}
\addtolength{\evensidemargin}{-1.0in}
\addtolength{\textwidth}{2.0in}
\addtolength{\topmargin}{-0.8in}
\addtolength{\textheight}{1.6in}
% ==============================================================================

% ================================= DOCUMENT ===================================
\begin{document}
    \renewcommand\marginsymbol[1][0pt]{%
  \tabto*{0cm}\makebox[-1cm][c]{$\mathbb{P}$}\tabto*{\TabPrevPos}}

\maketitle
\input{../intro.tex}

\section{Simple search algorithms}

Given the function \(f(x) = -x^4 + 2x^3 + 2x^2 - x\):

\subsection{Derivative}
What is its derivative \(f'(x)\) ?

\subsection{Plotting \marginsymbol}
Plot the function, and its gradient(derivative) from \(x=-2\) to \(x=3\).
Use python, wolfram alpha or another plotting tool of your choice.

\subsection{Gradient Ascent \marginsymbol}
\label{subsec:grada}
Maximize using gradient ascent.
You can try step size 0.1 and start somewhere in the range [-2, 3].
How does the choice of starting point and step size affect the algorithm's performance?
Is there a starting point where the algorithm would not even be able to find a local maximum?

\subsection{Exhaustive Search \marginsymbol}
\label{subsec:exhaust}
Assume that we are only interested in maxima of \(f(x)\),
where \(-2\leq x \leq 3\),
and \(x\) increases in steps of length 0.5 (\(\Delta x = 0.5\)).
Perform an exhaustive search to maximize \(f(x)\) and plot the result.

\subsection{Greedy Search and Hill Climbing}
In what way would greedy search and hill climbing differ for the maximization problem in Problem \ref{subsec:grada}?
Can you identify a starting position where the two algorithms might give different results?

\subsection{Possible improvements}
Gradient ascent, greedy search and hill climbing are quite similar, and are all based almost exclusively on exploitation.
Can you think of any additions to these algorithms in order to do more exploration?

\subsection{Exhaustive search vs. simulated annealing}
Which algorithm do you think is the most efficient at maximizing \(f(x)\) under the conditions in Problem \ref{subsec:exhaust};
exhaustive search or simulated annealing? Explain.

\input{../contact.tex}

\end{document}
% ==============================================================================
