% ================================== HEADER ====================================
\documentclass{article}           % Sets style/look of many things.
% \documentclass{report}          % part, chapters, front page etc.
\usepackage{exsheets}
\usepackage[utf8]{inputenc}       % Encoding of input files UTF-8
\usepackage[T1]{fontenc}
\usepackage[scaled]{beramono}     % Font
\usepackage{color}                % Color text
\usepackage{titlesec}             % Select alternative section titles
\usepackage{fancyvrb}
\usepackage{verbatim}             % Comment environment
\usepackage{listings}             % Format and render text/code etc.
\usepackage{minted}               % Much better syntax highlighting
\usepackage{float}                % Control of floating environment/figure
\usepackage{graphicx,  subfigure} % Better figures, graphics, units etc.
\usepackage{multicol}             % Multiple columns
\usepackage{amsmath}              % Math: Equation, split, align etc.
\usepackage{siunitx}              % SI units
\usepackage{mathtools}            % Different math tools to use with amsmath
\usepackage{amssymb}              % Math symbols
\usepackage[
    colorlinks,
    citecolor=black,              % I like links with standard black color
    filecolor=black,
    linkcolor=black,
    urlcolor=black
]{hyperref}                       % Links in TOC etc.
\usepackage[all]{hypcap}          % Better links to floating environment

\usepackage{tabto}
\newcommand\marginsymbol[1][0pt]{%
  \tabto*{0cm}\makebox[\dimexpr-1cm-#1\relax][r]{$\mathbb{P}$}\tabto*{\TabPrevPos}}

\renewcommand{\thesubsection}{\thesection.\alph{subsection}}
\title{\vspace{-2cm}INF3490/INF4490 Exercise Solutions - Week 2}
\author{Ole Herman S. Elgesem}
\date{\today}

% Removing paragraph indents is sometimes useful:
\setlength\parindent{0pt}

% Make margins smaller to fit more figures, tables etc on page: (optional)
\addtolength{\oddsidemargin}{-1.0in}
\addtolength{\evensidemargin}{-1.0in}
\addtolength{\textwidth}{2.0in}
\addtolength{\topmargin}{-0.8in}
\addtolength{\textheight}{1.6in}
% ==============================================================================

% ================================= DOCUMENT ===================================
\begin{document}
    \renewcommand\marginsymbol[1][0pt]{%
  \tabto*{0cm}\makebox[-1cm][c]{$\mathbb{P}$}\tabto*{\TabPrevPos}}

\maketitle
\(\mathbb{P}\) marks the programming exercises, we strongly recommend using
the python programming language for these. Exercises may be added/changed
after publishing.

\section{Representations}
Recall all the representations that have been presented. Which mutation and
recombination operators are compatible with which representations?\\

\emph{Answer:}

\begin{itemize}
    \item Binary representation
    \begin{itemize}
        \item Bit-flip mutation
        \item N-point and uniform crossover
    \end{itemize}
    \item Integer representation
    \begin{itemize}
        \item Random reset and creep mutation
        \item N-point and uniform crossover
    \end{itemize}
    \item Cardinal/enumerated/symbolic representations
    \begin{itemize}
        \item Random reset mutation
        \item N-point and uniform crossover
    \end{itemize}
    \item Real-valued/Continuous representation
    \begin{itemize}
        \item Uniform and Gaussian mutation
        \item N-point, discrete uniform and arithmetic crossover
    \end{itemize}
    \item Permutation representation
    \begin{itemize}
        \item Swap, insert, scramble and invert mutation
        \item Partially mapped, order, cycle and edge crossover
    \end{itemize}
    \item Tree representation
    \begin{itemize}
        \item Mutation by random replacement
        \item Subtree swap mutation
    \end{itemize}
\end{itemize}

\section{Bit flip mutation}
Given the binary chromosome with length 4, calculate the probability that no
bits, one bit and more than one bit will be flipped in a bit-flip mutation with
\(p_m = \frac{1}{4}\).\\

\emph{Answer:}

Probability of no mutation:
\begin{equation*}
    P(0) = \frac{3}{4} \frac{3}{4} \frac{3}{4} \frac{3}{4} = \frac{3^4}{4^4} \approx 32\%
\end{equation*}
\href{https://en.wikipedia.org/wiki/Binomial_distribution}{Binomial probability}:
\begin{equation}
    P = {n \choose k} p^k (1-p)^{n-k}
\end{equation}
\href{https://en.wikipedia.org/wiki/Binomial_coefficient}{Binomial coefficient}:
\begin{equation}
    {n \choose k} = \frac{n!}{k!(n-k)!}
\end{equation}
Wehere \(n\) is number of events, \(k\) is number of occurences wanted and \(p\) is the probability of a single event.\\

In this case we want exactly 1 mutation in 4 events, and the probability is 0.25:
\begin{align*}
    n &= 4\\
    k &= 1\\
    p &= p_m = 0.25\\
    P(1) &= {4 \choose 1} 0.25^1 (1-0.25)^{4-1}\\
    P(1) &= \frac{4!}{1!(4-1)!} 0.25^1 (1-0.25)^{4-1}\\
    P(1) &= \frac{4*3*2}{3*2} 0.25^1 (0.75)^{3}\\
    P(1) &= 4 * 0.25^1 (0.75)^{3}\\
    P(1) &= 0.75^3 \approx 42\%
\end{align*}
This makes intuitive sense and we could also arrive at this result without using the general formula.
For example, the probability of mutation (yes,no,no,no) is:
\begin{align*}
    P(1) &= \frac{1}{4} \frac{3}{4} \frac{3}{4} \frac{3}{4}\\
    P(1) &= 0.25 * 0.75^3
\end{align*}
And there are 4 variants, 4 places where the mutation can happen, so:
\begin{align*}
    P(1) &= 4 * (0.25 * 0.75^3)\\
    P(1) &= 0.75^3 \approx 42\%\\
\end{align*}
Finally, the probability of more than one mutation:
\begin{align*}
    P(2+) &= 1 - P(0) - P(1)\\
    P(2+) &\approx 1 - 32\% - 42\%\\
    P(2+) &\approx 26\%
\end{align*}

\section{Crossover \marginsymbol}
Given the sequences (2,4,7,1,3,6,8,9,5) and (5,9,8,6,2,4,1,3,7). Implement
these algorithms to create a new pair of solutions:
\renewcommand{\theenumi}{\alph{enumi}}
\begin{enumerate}
  \item Partially mapped crossover (PMX).
  \item Order crossover.
  \item Cycle crossover.
\end{enumerate}

\emph{Answer:}

\subsection{Partially mapped crossover}
\subsubsection{Output}
\begin{verbatim}
Parents:
[2, 4, 7, 1, 3, 6, 8, 9, 5]
[5, 9, 8, 6, 2, 4, 1, 3, 7]
Children:
[5, 9, 7, 1, 3, 6, 4, 2, 8]
[3, 1, 8, 6, 2, 4, 7, 9, 5]
\end{verbatim}
\subsubsection{Source code}
\inputminted{Python}{pmx.py}
\subsection{Order Crossover}
\subsubsection{Output}
\begin{verbatim}
Parents:
[2, 4, 7, 1, 3, 6, 8, 9, 5]
[5, 9, 8, 6, 2, 4, 1, 3, 7]
Children:
[9, 2, 4, 1, 3, 6, 8, 7, 5]
[7, 3, 8, 6, 2, 4, 1, 9, 5]
\end{verbatim}
\subsubsection{Source code}
\inputminted{Python}{order.py}
\subsection{Cycle Crossover}
\subsubsection{Output}
\begin{verbatim}
Parents:
[2, 4, 7, 1, 3, 6, 8, 9, 5]
[5, 9, 8, 6, 2, 4, 1, 3, 7]
Children:
[2, 4, 7, 1, 3, 6, 8, 9, 5]
[5, 9, 8, 6, 2, 4, 1, 3, 7]
\end{verbatim}
(These 2 parents have just one cycle, so the child will be the same as parent 1).
\subsubsection{Source code}
\inputminted{Python}{cycle.py}

\section*{Source code and files}
If you want to run the examples yourself they are avaialable through \href{https://github.com/olehermanse/INF3490-AI_Machine_Learning}{\textbf{GitHub}}.

\section*{Corrections and suggestions}
Corrections of grammar, language, notation or suggestions for improving these exercises are appreciated.
E-mail me at: \href{mailto:olehelg@uio.no}{\textbf{olehelg@uio.no}} or use
\href{https://github.com/olehermanse/INF3490-AI_Machine_Learning}{\textbf{GitHub}}
to submit an issue or create a pull request.
\end{document}
% ==============================================================================
