\input{../header.tex}
\title{\vspace{-2cm}INF3490/INF4490 Exercises - Evolutionary Computing}
\author{Eivind Samuelsen\input{../author_footnote.tex}}
\date{}

% Removing paragraph indents is sometimes useful:
\setlength\parindent{0pt}
% ==============================================================================

% ================================= DOCUMENT ===================================
\begin{document}
    \renewcommand\marginsymbol[1][0pt]{%
  \tabto*{0cm}\makebox[-1cm][c]{$\mathbb{P}$}\tabto*{\TabPrevPos}}

\maketitle
\input{../intro.tex}

\section{Evolution strategy(ES)} % 1
\subsection{} % a
A common variant of evolution strategies used for (local) search is the \((1 + 4)\) ES.
How would this differ from the \((1 + 1)\) ES in how the search space is explored?
How does this, and \((1 + \lambda)\) in general, compared to hill climbing and greedy search?
\subsection{} % b
What effect does an adaptive search strategy have on optimization performance?
\subsection{} % c
How would it affect the search if the strategy parameters were mutated after the solution parameters instead of before?
\section{ES Implementation} % 2
\subsection{\marginsymbol} % a
\label{subsec:w4e3a}
Ignoring mutation, and starting with the population \(\{1, 2, 3, 4\}\),
implement and run 3 generations of a \((4 + 8)\) ES maximizing \(g(x) = x\),
and observe what the end population looks like (use intermediary recombination).
\subsection{} % b
If a \((4, 8)\) ES had been used in Problem \ref{subsec:w4e3a},
what would the probability of the optimal solution \((x = 4)\) surviving the first generation have been?
\subsection{\marginsymbol} % c
Repeat Problem \ref{subsec:w4e3a} with an EP with \(q = 2\).
How do the two algorithms compare?
\section{Knapsack problem}
In a 0-1 \href{https://en.wikipedia.org/wiki/Knapsack_problem}{\textbf{knapsack problem}},
how could you implement a repair mutation to transform infeasible solutions into feasible ones
(i.e. make the sum of costs of the selected items go below the budget)?

\input{../contact.tex}
\end{document}
% ==============================================================================
