\input{../header.tex}
\title{\vspace{-2cm}INF3490/INF4490 Exercises - Support Vector Machines}
\author{Eivind Samuelsen\input{../author_footnote.tex}}
\date{}

% Removing paragraph indents is sometimes useful:
\setlength\parindent{0pt}
% ==============================================================================

% ================================= DOCUMENT ===================================
\begin{document}
    \renewcommand\marginsymbol[1][0pt]{%
  \tabto*{0cm}\makebox[-1cm][c]{$\mathbb{P}$}\tabto*{\TabPrevPos}}

\maketitle
\input{../intro.tex}

\section{SVM vs MLP}
What advantages and disadvantages are there to support vector machines (svm) versus multilayer perceptrons (mlp)?
What problems do they both suffer from?

\section{Kernel functions}
What is a kernel function?
Which are the most common kernel functions and roughly what kind of transformations do they correspond to?

\section{Soft Margins}
What two factors must be balanced when using an SVM with soft margin?

\section{Ensemble}
Try to come up with a few cases when using an ensemble of classifiers where it would be fine to just pick the most popular class, and where you would want to have the majority in favor of a single class or even full consensus.

\section{Principle Component Analysis}
What is the motivation behind principle component analysis?

\section{Covariance}
Work out the covariance between the x and y dimensions of the following 2-dimensional data set.
Describe what the results indicate about the data.

\begin{table}[H]
\centering
  \begin{tabular}{c|c|c|c|c|c}
    Index &  1 &  2 &  3 &  4 &  5 \\ \hline
    x     & 10 & 39 & 19 & 23 & 28 \\
    y     & 43 & 13 & 32 & 21 & 20 \\ \hline
  \end{tabular}
  \caption{Two-dimensional data set}
  \label{tab:cov}
\end{table}

\input{../contact.tex}
\end{document}
% ==============================================================================
